### 2017-01-04 robust-software.tex
Reorganizing "print usage" and "configurability".
1.  Combining requirement to log actions with requirement to print usage.
2.  Moving configurability to immediately follow command-line usage requirement.

Note: one practice \emph{not} included in the list below is the use of version control.
We think this is essential to building software,
and indeed to doing reproducible scientific work of any kind
\cite{wilson2014,wilson2016}.
However,
our focus in this short guide is on the software that is built,
not on the process used to build it.

@@

\section{Print usage information when launching from the command line that explains the software's features.}

Usage information provides the first line of help for both first-time and
experienced users of command-line applications.
Ideally, usage is a terse, informative command-line help message that
guides the user in the correct use of the software. Terseness is
important: usage that extends for multiple screens is a nuisance, especially when
printed to standard error instead of standard output (where it can
easily be paged).  Long output is also much less likely to be read{\ldots}

@@

\textbf{The syntax for running the program}: This includes the name of the
program and defines the relative
location of optional and required flags, arguments and values for execution.
Arguments in {[}square brackets{]}
are usually optional. An ellipsis (\ldots e.g. ``{[}OPTION{]}\ldots{}'')
indicates that more than one value can be provided.

\textbf{Description}: Similar to the README, the
description reminds users of the software's primary function.

\textbf{Most commonly used arguments, a description of each, and the
default values}: Not all arguments need to appear in the usage, but those most
commonly used should be listed. Users will rely on this for
quick reference when working with the software.

\textbf{Where to find more information}: Whether it's an email address, web
site, or manual page, there should be an indication where the user can go
to find out more about the software.

\textbf{Printed to standard output}: Usage should be displayed on standard output so that it can be piped into
\texttt{less}, searched with \texttt{grep}, or compared to the previous
version with \texttt{diff}.

\textbf{Exit with an appropriate exit code}: When usage is invoked by
providing incorrect arguments, the program should exit with a non-zero
code to indicate an error. However, when help is explicitly requested,
the software should not exit with an error.

@@

Providing parameters on the
command line increases the flexibility and usability of the program. You
may have determined early in development that 0.58 is an optimal cluster threshold for your
original dataset, but that doesn't mean it is best for every
case. Being able to change parameters on the fly to determine if and how
they change the results is important as your software gains more users,
as it facilitates exploratory analysis and parameter sweeping.
\textbf{Choose reasonable defaults where they exist},
and \textbf{set no defaults at all when there aren't any reasonable ones}.

You can set reasonable default values
as long as any command line arguments
override those values. Configuration files should be used in preference
to hard-coding the defaults directly. Configuration files can be in a standard
location, e.g. \texttt{.packagerc} in the user's home directory or
provided on the command line as an additional argument.
Only values that are unlikely to
change between runs belong in the config file, such as dependencies,
servers, version numbers, network drives, and any other defaults for
your lab or institution. 
Use a default configuration file and specify all other parameters on the
command line.

When the software starts, it should \textbf{echo all parameters and software
versions to standard out or a log file alongside the results}. This
feature supports greater reproducibility because any result can be
replicated with only the previous output files as reference.
Finally,
if the software is configurable in any way,
it should \textbf{check that all input values are in a reasonable range at startup}. 
Few things are as annoying as having a program announce after running for two hours
that it isn't going to save its results
because the requested directory doesn't exist.

>>>>

@@

\section{Tell the user what could be done and what was actually done.}

A robust program should
\textbf{print usage information when launching from the command line that explains the software's features}.
Usage information provides the first line of help for both first-time and
experienced users of command-line applications.
Ideally, usage is a terse, informative command-line help message that
guides the user in the correct use of the software. Terseness is
important: usage that extends for multiple screens is a nuisance, especially when
printed to standard error instead of standard output (where it can
easily be paged).  Long output is also much less likely to be read{\ldots}

@@

The key features of this usage message are:

\begin{itemize}

\item[\textbf{The syntax for running the program}] This includes the
  name of the program and defines the relative location of optional
  and required flags, arguments and values for execution.  Arguments
  in {[}square brackets{]} are usually optional. An ellipsis (\ldots
  e.g. ``{[}OPTION{]}\ldots{}'') indicates that more than one value
  can be provided.

\item[\textbf{Description}] Similar to the README, the description
  reminds users of the software's primary function.

\item[\textbf{Most commonly used arguments, a description of each, and
    the default values}] Not all arguments need to appear in the
  usage, but those most commonly used should be listed. Users will
  rely on this for quick reference when working with the software.

\item[\textbf{Where to find more information}] Whether it's an email
  address, web site, or manual page, there should be an indication
  where the user can go to find out more about the software.

\item[\textbf{Printed to standard output}] Usage should be displayed
  on standard output so that it can be piped into \texttt{less},
  searched with \texttt{grep}, or compared to the previous version
  with \texttt{diff}.

\item[\textbf{Exit with an appropriate exit code}] When usage is
  invoked by providing incorrect arguments, the program should exit
  with a non-zero code to indicate an error. However, when help is
  explicitly requested, the software should not exit with an error.

\end{itemize}

The usage message tells users what the program could do.
It is equally important for the program to tell users what it actually did.
Accordingly,
when the program starts, it should \textbf{echo all parameters and software
versions to standard out or a log file alongside the results}. This
feature supports greater reproducibility because any result can be
replicated with only the previous output files as reference.

(And as a corollary
if the software is configurable in any way,
it should \textbf{check that all input values are in a reasonable range at startup}. 
Few things are as annoying as having a program announce after running for two hours
that it isn't going to save its results
because the requested directory doesn't exist.)

Allowing parameters to be specific on the
command line increases the flexibility and usability of the program. You
may have determined early in development that 0.58 is an optimal cluster threshold for your
original dataset, but that doesn't mean it is best for every
case. Being able to change parameters on the fly to determine if and how
they change the results is important as your software gains more users,
as it facilitates exploratory analysis and parameter sweeping.

\textbf{Choose reasonable defaults where they exist},
and \textbf{set no defaults at all when there aren't any reasonable ones}.
You can set reasonable default values
as long as any command line arguments
override those values. Configuration files should be used in preference
to hard-coding the defaults directly. Configuration files can be in a standard
location, e.g. \texttt{.packagerc} in the user's home directory or
provided on the command line as an additional argument.
Only values that are unlikely to
change between runs belong in the config file, such as dependencies,
servers, version numbers, network drives, and any other defaults for
your lab or institution. 
Use a default configuration file and specify all other parameters on the
command line.

@@
